Labor Markets and Technological Change: Evidence from Electronic Health Records

### 2022-02-22 Paper/Paper.tex
Updates from Overleaf

Medicare Data on Provider Practice and Specialty is a database of physicians designed to record information about claim counts and specialty from 2009-2017. The relevant information included is physician NPI, primary and secondary specialties, patient counts in up to 12 different zip codes, and fraction of patients in different settings such as office, inpatient or emergency room. First, I make limitations to the sample of physicians based on specialty type. Primary care physicians are the most likely to be exposed to the administrative burden of EHRs in hospitals. Thus, I only include physicians whose primary specialty is hospitalist, internal medicine, pediatrics, general practice or family practice in at least one year (excluding physicians who have specialist listed for all other years). Since the focus of the analysis is physicians working in hospitals, I also drop physicians with less than 20\% of patients in a hospital setting in all years. 
The dependent variables are constructed using MD-PPAS claim and patient count information. Retirement is an indicator variable that is equal to 1 in the first year that all future claims in all zip codes are equal to 0. I do not observe retirement in 2017, and drop those who "retire" in 2016 since one year of zero claims does not necessarily mean retirement.

@@

CMS Shared Patient data records the number of patients who bill two entities under Medicaid in the same day. For example, if a Medicaid patient is referred to a specialist by a primary care physician, then those two physicians have a shared patient in common. I limit the entities by NPI tax-code to only include shared patients for physician-hospital pairs. I care specifically about physicians who have a close working relationship with hospitals, preferably those who do rounds within at least one hospital. To achieve this, I limit the qualifications for a sufficient physician-hospital pair. First, I only consider doctors who have a primary care role, excluding any specialists. The reason for this is that a primary care physician who does rounds in a hospital is likely to be primarily employed by that hospital (not brought in from a practice). Another reason for this is that primary care doctors interact directly with EHRs during checkups, unlike a surgeon. Second, I limit pairs to those which have a substantial number of patients billed together in the same day. This is to avoid including the potential physician who works in an office but happens to send a small number of patients to a nearby hospital. I consider a substantial number of patients to be at least 30 patients bill both the physician and the hospital in the same day in at least one year, \textcolor{red}{where the exact number chosen is explored in the appendix}. The physicians remaining in the data are likely doing a number of rounds physically inside hospitals, since it is unlikely that many patients will see a primary care doctor in a private office and then immediately go to a hospital in the same day.

Using hospital NPI, I link the physician-hospital pairs to the American Hospital Association survey, which contains information on hospital-level EHR use and other characteristics. I then aggregate to the physician level, where physician EHR use is measured as the percentage of their hospitals which use an EHR in a given year. Additional physician-level characteristics, such as years of experience, come from Physician Compare.

>>>>

Medicare Data on Provider Practice and Specialty is a database of physicians designed to record information about claim counts and specialty from 2009-2017. The relevant information included is physician NPI, primary and secondary specialties, patient counts in up to 12 different zip codes, and fraction of patients in different settings such as office, inpatient or emergency room. First, I make limitations to the sample of physicians based on specialty type. Primary care physicians are the most likely to be exposed to the administrative burden of EHRs in hospitals. Thus, I only include physicians whose primary specialty is hospitalist, internal medicine, pediatrics, general practice or family practice in at least one year (excluding physicians who have specialist listed for all other years). Since the focus of the analysis is physicians working in hospitals, I also drop physicians with less than 20\% of patients in a hospital setting in all years. 
The dependent variables are constructed using MD-PPAS claim and patient count information. Retirement is an indicator variable that is equal to 1 in the first year that all future claims in all zip codes are equal to 0. I do not observe retirement in 2017, and drop those who "retire" in 2016 since one year of zero claims does not necessarily mean retirement. The average age of the sample is 51.5, indicating a large number of older physicians, where the average age of those who retire at some point in the sample is 52.5. This indicates a large older population in the sample. \textcolor{red}{maybe show the distribution of ages by those who retire and those who don't.}

\textcolor{red}{paragraph about switching setting variables}

\textcolor{red}{paragraph about patient count variable}

@@

CMS Shared Patient data records the number of patients who bill two entities under Medicaid in the same day. For example, if a Medicaid patient is referred to a specialist by a primary care physician and the specialist is seen the same day, then those two physicians have a shared patient in common. I utilize this data to find physicians' EHR exposure based on the hospitals they work with, which defines treatment variables for the analysis. I limit the entities by NPI tax-code to only include shared patients for physician-hospital pairs. I care specifically about physicians who have a close working relationship with hospitals, preferably those who do rounds within at least one hospital. The types of physicians are already limited to primary care physicians, but there still may be some office-based primary care physicians in the shared patient data. Therefore, I limit the sample of pairs based on a threshold of same day patients which depends on the number of years the physician appears in the data. A physician is dropped if their total number of same day billings with a hospital is less than 30 times the number of years the pair appears in the data. The reason for this is to avoid dropping physicians who drop out of the data for many years, as this may indicate retirement. \textcolor{red}{I explore other thresholds in the appendix.}

Using hospital NPI, I link the physician-hospital pairs to the American Hospital Association survey, which contains information on hospital-level EHR use and other characteristics. Then I record the first year a physician is exposed to an EHR based on the hospitals they share patients with. I create two measures of EHR exposure. The first is exposure to any EHR in any hospital, and the second is exposure to EHR at a "low-integration" hospital (a hospital where physicians have a low measure of vertical integration, as defined in \textcolor{red}{add cite}). I then aggregate to the physician level.

\textcolor{red}{write about the discrepancy in years}

\subsection{Summary Statistics}
